Challenging Fuel Cycle Modeling Assumptions:\\Facility and Time Step Discretization Effects

### 2016-05-04 paper.tex
mods to format/style for arxiv.org submission

%%%%%%%% cover info %%%%%%%%%%%%%%%
\title{Challenging Fuel Cycle Modeling Assumptions: Facility and Time Step Discretization Effects}
\author{Robert~W.~Carlsen, Paul~P.H.~Wilson}

\institute{University of Wisconsin - Madison, Department of Nuclear Engineering and Engineering Physics}
\instituteaddress{1415 Engineering Drive, Madison, WI 53706}

@@

\myabstract{
Due to the diversity of fuel cycle simulator modeling assumptions, direct
comparison and benchmarking can be difficult.  In 2012 the Organisation for
Economic Co-operation and Development (OECD) completed a benchmark study that
is perhaps the most complete published comparison performed.  Despite this,
various results from the simulators were often significantly different because of
inconsistencies in modeling decisions involving reprocessing strategies,
refueling behavior, reactor end-of-life handling, etc.  This work identifies
and quantifies the effects of selected modeling choices that may sometimes be
taken for granted in the fuel cycle simulation domain.  Four scenarios are
compared using combinations of either fleet-based or individually modeled
reactors with either monthly or quarterly (3-month) time steps.  The scenarios
approximate a transition from the current U.S. once-through light water
reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle. The Cyclus
fuel cycle simulator's plug-in facility capability along with its market-like
dynamic material routing allow it to be used as a level playing field for
comparing the scenarios.  When under supply-constraint pressure, the four
cases exhibit noticeably different behavior.  Fleet-based modeling is more
efficient in supply-constrained environments at the expense of losing insight
on issues such as realistically suboptimal fuel distribution and challenges in
reactor refueling cycle staggering.  Finer-grained time steps also enable more
efficient material use in supply-constrained environments resulting in much lower
standing inventories of separated Pu.  Large simulations with fleet-based
reactors run much more quickly than their individual reactor counterparts.
Gaining a better understanding of how these and other modeling choices affect
fuel cycle dynamics will enable making more deliberate decisions with respect
to trade-offs such as computational investment vs.  realism.
}

>>>>

%%%%%%%% cover info %%%%%%%%%%%%%%%
\title{Challenging Fuel Cycle Modeling Assumptions:\\Facility and Time Step Discretization Effects}
\author{Robert~W.~Carlsen (rwcarlsen@gmail.com), Paul~P.H.~Wilson}

\institute{University of Wisconsin -- Madison, Department of Nuclear Engineering and Engineering Physics, Madison, WI 53706}
\instituteaddress{x}

@@

\myabstract{
\noindent Due to the diversity of fuel cycle simulator modeling assumptions, direct
comparison and benchmarking can be difficult.  In 2012 the Organisation for
Economic Co-operation and Development (OECD) completed a benchmark study that
is perhaps the most complete published comparison performed.  Despite this,
various results from the simulators were often significantly different because of
inconsistencies in modeling decisions involving reprocessing strategies,
refueling behavior, reactor end-of-life handling, etc.  This work identifies
and quantifies the effects of selected modeling choices that may sometimes be
taken for granted in the fuel cycle simulation domain.  Four scenarios are
compared using combinations of either fleet-based or individually modeled
reactors with either monthly or quarterly (3-month) time steps.  The scenarios
approximate a transition from the current U.S. once-through light water
reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle. The Cyclus
fuel cycle simulator's plug-in facility capability along with its market-like
dynamic material routing allow it to be used as a level playing field for
comparing the scenarios.  When under supply-constraint pressure, the four
cases exhibit noticeably different behavior.  Fleet-based modeling is more
efficient in supply-constrained environments at the expense of losing insight
on issues such as realistically suboptimal fuel distribution and challenges in
reactor refueling cycle staggering.  Finer-grained time steps also enable more
efficient material use in supply-constrained environments resulting in much lower
standing inventories of separated Pu.  Large simulations with fleet-based
reactors run much more quickly than their individual reactor counterparts.
Gaining a better understanding of how these and other modeling choices affect
fuel cycle dynamics will enable making more deliberate decisions with respect
to trade-offs such as computational investment vs.  realism.
}
